Canonical rings of $\bq$-divisors on Minimal Rational Surfaces

### 2015-07-26 surfaces-write-up/surface-write-up.tex
responded mostly to todos in hirzebruch section, replaced reals by rationls

\begin{defn}
\label{defn:lower-approximation}
If $\alpha \in \br$, then a rational number $\frac{c}{k} \leq \alpha$
(written in reduced form with $c \in \bz, k \in \bn$) is a
\textbf{best lower approximation} to $\alpha$ if there does not
exist $\frac{c'}{k'}\in \mathbf{Q}$ such that $0 < k'< k$ and
$\frac{c}{k} \le \frac{c'}{k'} \le \alpha$. 
\end{defn}

@@

\noindent
Suppose $e_1, \ldots, e_t \in \Sigma$ with $e_i = (\pdeg_i, c_1^i,
\ldots, c_n^i)$ are a set of {\bf extremal rays} of $\Sigma$,
i.e.~ $\Sigma$ is contained in the $R_{\geq 0}$ span of
$e_1, \ldots, e_t$.

@@

\begin{proof}
By assumption, $\sigma \in \Sigma$ can be written as $\sigma = \sum_
{i = 1}^{t} r_i e_i$ with $r_i \in \br$. Let $\fr(r) := r - \lfloor r
\rfloor$ denote the fractional part of $r$. Let $\lambda = \sum_{i = 1}
^{t} \fr(r_i)$. Whence, we can write $\sigma = \lambda + \sum_{i = 1}
^{t} \lfloor r_i \rfloor e_i.$ Consequently, $\sigma$ lies in the
$\bz_{\geq 0}$ span of $\lambda, e_1, \ldots, e_t$, which all have
first coordinate less than $\sum_{i=1}^{t} r_i$. Ergo, $\Sigma$ is
generated by elements whose first coordinate is less than
$\sum_{i = 1}^{t} r_i$.
\end{proof}

@@

\begin{convention}
Let $[x_0, \ldots, x_m]$ be coordinates for $\bp^m$. Let
$\vec{v} = (v_0, \ldots, v_m) \in \br^{m + 1}$. Then write
\[
	x^{\vec{v}} := \prod_{i = 0}^{m} x_i^{v_i}
\]
\end{convention}

\begin{defn}
\label{defn:vec-sum}
For $\vec{v} \in \br^n$, denote $\deg \vec{v} := \sum_{i = 0}
^n v_i$.
Let 
\begin{align*}
	\mss_i := \left \{\vec{v} \in \bz_{\geq 0}^{m + 1} \; : \;
\deg \vec v = c_i \right\}.	
\end{align*}
\end{defn}

\begin{defn}
\label{defn:vec-order}
Let $\vec{v}, \vec{w} \in \br^n$. Let $i \in \{1,\ldots, n\}$
be the biggest index such that $v_i$ is nonzero
and $j \in \{1,\ldots, n\}$ be the smallest index such that $w_j$ is
nonzero. Define $\vec{v} \prec \vec{w}$ if $i \leq j$.
\end{defn}

@@

\begin{defn}\label{def:ell_i}
Let $D = \sum_{i=1}^n \alpha_i D_i$ where $\alpha_i = \frac{c_i}{k_i}$. \todo{Peter: better notation for fraction?}  Then notate
\[\ell_i = \lcm_{j\ne i}(k_i).\]
\end{defn}

\begin{lem} \label{lem:proj-generators}
$R_D$ is generated in degrees up to $\sum_{i=1}^n \ell_i a_i.$
% and relations in degree at most  .\todo{insert numbers}
\end{lem}
\begin{proof}
Let 
\begin{align}\label{eqn:Sigma-def}
	\Sigma = \left \{(d, c_1, \ldots, c_n) \in \bz^{n+1} : c_i \geq - d
\alpha_i, \; 1 \leq i \leq n, \text{ and } \sum_{i=1}^{n} \ell_i \bida_i = 0
\right \}.
\end{align}
\todo{Possibly take this definition out of the proof, and move it into the setup, since we want to reference it later.}
We note that $\Sigma$ has extremal rays given by the lattice points 
\begin{equation}\label{defn:e-i-proj}
	e_i = \left(\ell_i \bida_i, - \alpha_1 \ell_i \bida_i, \ldots
-\alpha_{i-1} \ell_i \bida_i, \ell_i \sum_{j\ne i} \alpha_j \bida_j,
-\alpha_{i+1} \ell_i \bida_i, \ldots, -\alpha_n, \ell_i \bida_i \right)
\end{equation}
for each $i\in \{1, \ldots n\}$.
Therefore, applying Proposition ~\ref{prop:cone-generation}, we see $R_D$ is generated in degrees less than
\[
	\sum_{i=1}^n \ell_i a_i.
\]
\end{proof}

@@

\begin{lem}
\label{lem:proj-relations-psi}
$\ker(\psi)$ is generated in degrees up to
\begin{align}
\label{eqn:proj-relation-degree}
	\frac{\max_{1\le i \le n}(\bida_i)}{\deg(D)} +  \sum_{i=1}^n \ell_i a_i.
\end{align}

@@

\noindent
Then we see $\Sigma_i \subseteq \delta_i + \newspan_{\br_{\geq 0}}
(e_1, \ldots, e_n)$ with $e_i$ as defined in Equation
~\ref{defn:e-i-proj}. Therefore, we can write any element of
$\Sigma_i$ as
\[
	\delta_i + \sum_{j=1}^n r_j e_j
\]
where $r_j \in \mathbb{R}$ for each $j$.

Whenever some $r_j \ge 1$, we can generate the relation $(z_i -
\beta_i)\prod_{j=1}^n z_j^{c_j} = \epsilon_j \cdot h,$ where
$\epsilon_j$ was defined in Equation ~\ref{eqn:epsilon-def-proj} and
$h \in \Sigma_i$. Therefore, $\ker(\psi)$ is generated in degrees
less than
\[
	\frac{\bida_i}{\deg(D)} + \sum_{i=1}^n \ell_i a_i.
\]
\end{proof}

@@

By ~\ref{lem:bound-ker-chi}, $\ker(\chi)$ is generated in degrees 
up to $2\sum_{i=1}^n \ell_i a_i$ and by Lemma
~\ref{lem:proj-relations-psi}, $\ker(\psi),$ is generated in
degrees up to $\frac{\bida_i}{\deg(D)} + \sum_{i=1}^n \ell_i a_i$. 
Consequently, Lemma ~\ref{lem:composite-map} implies that $\ker(\phi)$
is generated in degrees less than
\[
	\max \left(2 \sum_{i=1}^n \ell_i a_i, \frac{\max_{1\le i \le n}
	(\bida_i)}{\deg(D)} + \sum_{i=1}^n \ell_i a_i \right).
\]
\end{proof}

@@

We can talk about sections of $\hirz_m$ in terms of polynomials in $u,v, w,x$; furthermore we can talk about bi-degrees of a monomial $u^a v^{b} w^c x^d$ given by $(a + b + mc, c + d)$.  Sections on open subsets of $\hirz_m$ look like polynomials with each monomial having the same bidegrees, and rational sections have both bidegrees 0. \todo{Peter @ Aaron: can you add your exact sequence argument for why this is true?}

Why can write $\bq$-divisors of $\hirz_m$ as $D = \sum_{i = 1}^n \alpha_i D_i \in \mathbb{Q} \otimes \di (\hirz_m)$, where each $D_i = V(f_i)$ for some polynomial $f_i$, with each monomial having the same bidegrees.  Since sections on $H^0(dD)$ must be rational sections, they must have both bidegrees non-negative, meaning that both bidegrees of $D$ must be non-negative for $R_D$ to be non-trivial.

For the remainder of this section we will assume $D_1, D_2, D_3,$ and $D_4$ are distinct divisors with bi-degrees $(1,0)$
, $(1,0)$, $(0,1)$, and $(0,1)$ respectively. Let $D_i = V(f_i)$ where
$f_i \in \sco(\bida_i, \bidb_i)$.  
Then, $f_1$ and $f_2$ are independent linear
polynomials in $x_0$ and $x_1$ and $f_3$ and $f_4$ are independent
linear polynomials in $y_0$ and $y_1$. \todo{Talk a little about Hirzebruch
surfaces before this paragraph; especially, describe what it's sections look
like. \\
Peter: Fixed?
}

\begin{defn}
Define 
\begin{equation}
\label{eqn:define-T=(D)}
	\Te(D) = \left\{i \in \{1, \ldots, n\}: \bida_i \sum_{k=1}^n \bidb_k 
\alpha_k = \bidb_i \sum_{k=1}^n \bida_k \alpha_k \right\},
\end{equation}

@@

\begin{defn}\label{def:ell_ij}
Let $D = \sum_{i=1}^n \alpha_i D_i$ where $\alpha_i = \frac{c_i}{k_i}$. \todo{Peter: better notation for fraction?}  Then notate
\[\ell_{i,j} = \lcm_{h\ne i,j}(k_h).\]
\end{defn}

\begin{lem}
\label{lem:hirz-generators}
$R_D$ is generated in degree at most
\begin{equation}\label{eqn:def-sigma}
	\rho = \sum_{i \in \Te(D)} \gcd(\bida_i, \bidb_i) \ell_i + \sum_{\substack
	{i \in \Tp(D) \\ j\in \Tm(D)}} (\bida_i \bidb_j - \bida_j \bidb_i)
	\ell	_{i,j}.
\end{equation}
\end{lem}

\begin{proof}
%Let $(\bida_i, \bidb_i)$ be the bi-degrees of $f_i$. 
Suppose $g
\in (R_D)_d$.
 Then 
\[
	g = \sum_{i = 1}^n {f_i}^{c_i}
\]

\noindent
for some $c_i \ge - \alpha_i d$ such that $\sum_{i=1}^n c_i \bida_i = 0$
and $\sum_{i=1}^n c_i \bidb_i = 0$. We can view $g$ inside the
lattice 
\begin{align*}
	\Sigma = \{ (d, c_1, \ldots, c_n) \in \bz_{\geq 0}^{n+1} : c_i \ge - d
\alpha_i \text{ for all }i \text{ and }\sum_{i=1}^n c_i \bida_i = \sum_{i =
1}^n c_i \bidb_i = 0 \}.	
\end{align*}

@@

Let $S$ be the $\mathbb{R}_{\ge 0}$ span of $e_1, \ldots e_n$.
We can intersect $S$ with the hyperplane $H$ given by
$\sum_{i=1}^n \bida_i X_i = 0$  to get a subspace $S' \subseteq S$.  Then, as long as each extremal rays of $S'$ contains an element of the lattice $\Sigma$, those representatives must in fact the extremal rays of $\Sigma'$.

The extremal rays of $S'$ can be represented by points lying only on the edges of 
$\overline{e_i e_j}$
and are given by the $e_i$'s
contained in $H$ and the intersection points $e_{i, j}$ of $H$
together with the edges $\overline{e_i e_j}$ of the simplex between
distinct $e_i, e_j$ not contained in $H$.
\todo{more detail above;  Peter: resolved?}
The following are elements of $R_D$ correspond to lattice point on each of the extremal rays listed above:

For $i \in \Te(D)$, define $\epsilon_i \in R_D$ in degree $d_i = \ell_i \gcd(a_i, b_i)$ by 
\todo{below section needs to be simplified, maybe
by introducing new variables; 
Peter: I replaced some terms with degrees $d_i$ and $d_{i,j}$}
\begin{equation}\label{eqn:epsilon_i}
	\epsilon_i := \left(\prod_{k \ne i} f_k^{-\alpha_k d_i}\right)
	f_i^{ \frac{d_i}{\bida_i + \bidb_i}\sum_{k \ne i}
	\alpha_k (\bida_k + \bidb_k)}.
\end{equation}

@@

\noindent
and
\[
	s_2 = -\bidb_i \sum_{k \ne i, j} \alpha_k \bida_k + \bida_i \sum_{k \ne i, j}
	\alpha_k \bidb_k.
\]
\todo{Aaron: Why are $\epsilon_{i,j}$ defined here instead of $e_{i,j}$. We don't use the $\epsilon$ here, and we do use the $e$ here. \\
Peter: I have a slight leaning towards this notation, because the algebraic object is use more later (I think, although I'll check to make sure); also, I think writing out the ordered tuple would actually be harder, because there is no product notation, and it can be hard to denote the cordinates.}
A quick calculation deduces that these $\epsilon_i$'s and $\epsilon_{i,j}$'s lie on the extremal rays represented by the $e_i$'s and $e_{i,j}$'s, respectively.
\begin{rem}
For the remainder of this section replace the $e_i$'s and $e_{i,j}$'s with a (possibly) different representative of it's extremal ray say that their lattice points correspond to the $\epsilon_i$'s and $\epsilon_{i,j}$'s, respectively.
\end{rem}\todo{Peter: I don't like having a remark inside of a proof.  Also, it feels like this could reasonably be replaced with a better constructed/organized proof, so that $e_i$'s are just defined this way initially; let me know if you see a good way to do so...  As for Aaron's comment above, we could do something like that, except we already defined $e_i$'s to be something different, which might be equally/more confusing.}

Thus,
Proposition ~\ref{prop:cone-generation} implies that 
$R_D$ is generated in 
degrees less than the sum of the degrees of the $\epsilon_i$'s and $\epsilon_{i,j}$'s, which is
\[
	\rho = \sum_{i\in \Te(D)} \gcd(\bida_i, \bidb_i)\ell_i + \sum_{\substack{
	i \in \Tp(D) \\	j \in \Tm(D)}} (\bida_i \bidb_j- \bida_j \bidb_i)\ell_{i,j}.
\]
\todo{Aaron: Do you use $\sigma$ later? We are usually reserving $\sigma$ for a
lattice point. Perhaps something like $\delta$ would be more appropriate if you
want to refer to it later. Also, you should number if if you want to refer to
it later. In general, it's bad style to define things in the proofs. This could
have been defined prior to the theorem statement as well, which would simplify
the statement. We should also then clarify that $\sigma$ depends on $D$.  \\
Peter: $\sigma$ replaced by $\rho$. It's also defined now in the lemma statement rather than its proof.  
}

\end{proof}

@@

\begin{lem}
$\ker(\psi)$ is generated in degrees less than
\[
	\tau : = \rho
	+ \max \left(\max_{i\in \Te}(\ell_i \gcd(a_i, b_i)), \max_{\substack{
	i \in	\Tp \\ j\in \Tm}} (\ell_{i,j} (\bida_i \bidb_j - \bida_j \bidb_i))
	\right).
\]
\end{lem}

\begin{proof}
We first claim that there exist $\beta_1, \ldots, \beta_n\in \bk[z_1, z_2, z_3, z_4]$
 such that $\ker(\psi)$ has relations of the form
\begin{equation}
\label{eqn:hirz-relations-psi}
	(z_i - \beta_i)\prod_{j=1}^n {z_j}^{c_{j}}
\end{equation}

@@

Let $\omega\in \Omega'$ correspond to the monomial $y_i \prod_{j=1}^n
{y_j}^{c_j}$.  Then we can write $\omega$ as a sum of $r_j e_j$'s for $j\in
\Te$ and $r_{j,k}e_{j,k}$ for $j\in \Tp, k\in \Tm$.  
For convenience, let $e_j$ (when it exists) be in degree $d_j$ with $j^{th}$
component $-\alpha_j d_j + \kappa_j$ for some $\kappa_j \in \br$ and let $e_{j,k}$ (when it exists) lie in
degrees $d_{j,k}$ with $j^{th}$ component $-\alpha_j d_{j,k} + \kappa_{j,k}'$ and $k^{th}$
component $-\alpha_j d_{j,k} + \kappa_{j,k}''$ for $\kappa_{j,k}', \kappa_{j,k}'' \in \br$
Then a careful computation deduces that 
%if $i\in \Te$, $i\in \Tp$, and $i\in \Tm$, then
\begin{equation}\label{eqn:sigma-in-terms-of-lambda/E}
	\sigma = \lambda + E
\end{equation}
where $\lambda \in \Sigma$ and
\begin{equation}\label{eqn:hirz-E-translation}
	E = \begin{cases}
	\frac{1}{\kappa_i} e_i  	&  \mbox{ if }i\in \Te\\
	\sum_{j\in \Tm} r_{i,j} e_{i,j}: r_{i,j}\in \br, \sum_{j\in \Tm} r_{i,j}\kappa_{i,j}' = 1 & \mbox{ if } i\in \Tp\\
	\sum_{j\in \Tp} r_{j,i} e_{j,i}: r_{j,i} \in \br, \sum_{j\in \Tm} r_{j,i} \kappa_{i,j}'' = 1 	& \mbox{ if } i\in \Tm\\
	\end{cases}.
\end{equation}

\noindent
From this, we can deduce that
\[
	\deg(E) \le \begin{cases}
	\ell_i \gcd(\bida_i, \bidb_i)	&\mbox{ if } i \in \Te \\
	\max_{j \in \Tm} \bigl(\ell_{i, j} (\bida_i \bidb_j - \bida_j \bidb_i)\bigr)
	&\mbox{ if } i \in \Tp \\
	\max_{j \in \Tp} \bigl(\ell_{j, i} (\bida_j \bidb_ i - \bida_i \bidb_j) \bigr)
	&\mbox{ if } i \in \Tm \end{cases}.
\]
Let $M$ be the set of monomials terms of $\beta_i$.  Then for $\mu\in M$, we
can further compute that the lattice point, $\sigma_\mu$, corresponding to $\mu \prod_{j=1}^n
{f_i}^{c_i}$ is of the form
\[
	\sigma_\mu = \lambda + E_\mu
\]
where $\lambda \in \Sigma$ is the same as in Equation \ref{eqn:sigma-in-terms-of-lambda/E} and $E_\mu$ is defined in a similar fashion as $E$ in Equation \ref{eqn:hirz-E-translation} (in particular, they lie in the same degree).

Let $F\in R_D$ corresponding to the lattice point $E$ and let $F_\mu$ correspond to the lattice point $E_\mu$.  
Then $F - \sum_{\mu \in M} F_\mu = 0$ induces a
relation in degree at most
\[
	\rho + \deg(F)
\]

>>>>

\begin{defn}
\label{defn:lower-approximation}
If $\alpha \in \bq$, then a rational number $\frac{c}{k} \leq \alpha$
(written in reduced form with $c \in \bz, k \in \bn$) is a
\textbf{best lower approximation} to $\alpha$ if there does not
exist $\frac{c'}{k'}\in \mathbf{Q}$ such that $0 < k'< k$ and
$\frac{c}{k} \le \frac{c'}{k'} \le \alpha$. 
\end{defn}

@@

\noindent
Suppose $e_1, \ldots, e_t \in \Sigma$ with $e_i = (\pdeg_i, c_1^i,
\ldots, c_n^i)$ are a set of {\bf extremal rays} of $\Sigma$,
i.e.~ $\Sigma$ is contained in the $\bq_{\geq 0}$ span of
$e_1, \ldots, e_t$.

@@

\begin{proof}
By assumption, $\sigma \in \Sigma$ can be written as $\sigma = \sum_
{i = 1}^{t} r_i e_i$ with $r_i \in \bq$. Let $\fr(r) := r - \lfloor r
\rfloor$ denote the fractional part of $r$. Let $\lambda = \sum_{i = 1}
^{t} \fr(r_i)$. Whence, we can write $\sigma = \lambda + \sum_{i = 1}
^{t} \lfloor r_i \rfloor e_i.$ Consequently, $\sigma$ lies in the
$\bz_{\geq 0}$ span of $\lambda, e_1, \ldots, e_t$, which all have
first coordinate less than $\sum_{i=1}^{t} r_i$. Ergo, $\Sigma$ is
generated by elements whose first coordinate is less than
$\sum_{i = 1}^{t} r_i$.
\end{proof}

@@

\begin{convention}
Let $[x_0, \ldots, x_m]$ be coordinates for $\bp^m$. Let
$\vec{v} = (v_0, \ldots, v_m) \in \bq^{m + 1}$. Then write
\[
	x^{\vec{v}} := \prod_{i = 0}^{m} x_i^{v_i}
\]
\end{convention}

\begin{defn}
\label{defn:vec-sum}
For $\vec{v} \in \bq^n$, denote $\deg \vec{v} := \sum_{i = 0}
^n v_i$.
Let 
\begin{align*}
	\mss_i := \left \{\vec{v} \in \bz_{\geq 0}^{m + 1} \; : \;
\deg \vec v = c_i \right\}.	
\end{align*}
\end{defn}

\begin{defn}
\label{defn:vec-order}
Let $\vec{v}, \vec{w} \in \bq^n$. Let $i \in \{1,\ldots, n\}$
be the biggest index such that $v_i$ is nonzero
and $j \in \{1,\ldots, n\}$ be the smallest index such that $w_j$ is
nonzero. Define $\vec{v} \prec \vec{w}$ if $i \leq j$.
\end{defn}

@@

\begin{defn}\label{def:ell_i}
Let $D = \sum_{i=1}^n \alpha_i D_i$ where $\alpha_i = \frac{c_i}{k_i}$. \todo{Peter: better notation for fraction? Aaron: This seems fine to me. Are you saying you want a better notation, or asking if this is a better notation?}  Then notate
\[\ell_i = \lcm_{j\ne i}(k_i).\]
\end{defn}

\begin{lem} \label{lem:proj-generators}
$R_D$ is generated in degrees up to $\sum_{i=1}^n \ell_i a_i.$

\end{lem}
\begin{proof}
Let 
\begin{align}\label{eqn:Sigma-def}
	\Sigma = \left \{(d, c_1, \ldots, c_n) \in \bz^{n+1} : c_i \geq - d
\alpha_i, \; 1 \leq i \leq n, \text{ and } \sum_{i=1}^{n} \ell_i \bida_i = 0
\right \}.
\end{align}
\todo{Aaron: Possibly take this definition out of the proof, and move it into the setup, since we want to reference it later.}
We note that $\Sigma$ has extremal rays given by the lattice points 
\begin{equation}\label{defn:e-i-proj}
	e_i = \left(\ell_i \bida_i, - \alpha_1 \ell_i \bida_i, \ldots
-\alpha_{i-1} \ell_i \bida_i, \ell_i \sum_{j\ne i} \alpha_j \bida_j,
-\alpha_{i+1} \ell_i \bida_i, \ldots, -\alpha_n, \ell_i \bida_i \right)
\end{equation}
for each $i\in \{1, \ldots n\}$.
Therefore, applying Proposition ~\ref{prop:cone-generation}, we see $R_D$ is generated in degrees less than
\[
	\sum_{i=1}^n \ell_i a_i.
\]
\end{proof}

@@

\begin{lem}
\label{lem:proj-relations-psi}
$\ker \psi$ is generated in degrees up to
\begin{align}
\label{eqn:proj-relation-degree}
	\frac{\max_{1\le i \le n}(\bida_i)}{\deg(D)} +  \sum_{i=1}^n \ell_i a_i.
\end{align}

@@

\noindent
Then we see $\Sigma_i \subseteq \delta_i + \newspan_{\bq_{\geq 0}}
(e_1, \ldots, e_n)$ with $e_i$ as defined in Equation
~\ref{defn:e-i-proj}. Therefore, we can write any element of
$\Sigma_i$ as
\[
	\delta_i + \sum_{j=1}^n r_j e_j
\]
where $r_j \in \mathbb{R}$ for each $j$.

Whenever some $r_j \ge 1$, we can generate the relation $(z_i -
\beta_i)\prod_{j=1}^n z_j^{c_j} = \epsilon_j \cdot h,$ where
$\epsilon_j$ was defined in Equation ~\ref{eqn:epsilon-def-proj} and
$h \in \Sigma_i$. Therefore, $\ker \psi$ is generated in degrees
less than
\[
	\frac{\bida_i}{\deg(D)} + \sum_{i=1}^n \ell_i a_i.
\]
\end{proof}

@@

By ~\ref{lem:bound-ker-chi}, $\ker(\chi)$ is generated in degrees 
up to $2\sum_{i=1}^n \ell_i a_i$ and by Lemma
~\ref{lem:proj-relations-psi}, $\ker \psi,$ is generated in
degrees up to $\frac{\bida_i}{\deg(D)} + \sum_{i=1}^n \ell_i a_i$. 
Consequently, Lemma ~\ref{lem:composite-map} implies that $\ker \phi$
is generated in degrees less than
\[
	\max \left(2 \sum_{i=1}^n \ell_i a_i, \frac{\max_{1\le i \le n}
	(\bida_i)}{\deg(D)} + \sum_{i=1}^n \ell_i a_i \right).
\]
\end{proof}

@@

Sections of a line bundle $\mathscr L$ on $\hirz_m$ can be written as
polynomials in $u,v, w,x$.
Furthermore we define the bi-degree of a monomial $u^a v^{b} w^c x^d$
on $\hirz_n$ to be
$(a + b + mc, c + d)$.  Rational section of $\hirz_m$ can be written as
rational functions where each monomial has bi-degree 0. To see this, 
observe
$\Pic(\hirz_m) \cong \bz \times \bz$, where the class of a line bundle
in $\Pic(\hirz_m)$
is determined by its bi-degree, as follows from the excision exact sequence
for class groups.

For the remainder of this section we will assume $D_1, D_2, D_3,$ and $D_4$ are
distinct divisors with bi-degrees $(1,0)$
, $(1,0)$, $(0,1)$, and $(0,1)$ respectively. Let $D_i = V(f_i)$ where
$f_i \in \sco(\bida_i, \bidb_i)$.  
Then, $f_1$ and $f_2$ are independent linear
polynomials in $x_0$ and $x_1$ and $f_3$ and $f_4$ are independent
linear polynomials in $y_0$ and $y_1$.
\begin{defn}
Define 
\begin{equation}
\label{eqn:define-T=(D)}
	\Te(D) = \left\{i \in \{1, \ldots, n\}: \bida_i \sum_{k=1}^n \bidb_k 
\alpha_k = \bidb_i \sum_{k=1}^n \bida_k \alpha_k \right\},
\end{equation}

@@

\begin{defn}\label{def:ell_ij}
Let $D = \sum_{i=1}^n \alpha_i D_i$ where $\alpha_i = \frac{c_i}{k_i}$.
\todo{Peter: better notation for fraction? Aaron: I like this notation}  Then notate
\[\ell_{i,j} = \lcm_{h\ne i,j}(k_h).\]
\end{defn}

\begin{lem}
\label{lem:hirz-generators}
$R_D$ is generated in degree at most
\begin{equation}\label{eqn:def-sigma}
	\rho := \sum_{i \in \Te(D)} \gcd(\bida_i, \bidb_i) \ell_i + \sum_{\substack
	{i \in \Tp(D) \\ j\in \Tm(D)}} (\bida_i \bidb_j - \bida_j \bidb_i)
	\ell	_{i,j}.
\end{equation}
\end{lem}

\begin{proof}
%Let $(\bida_i, \bidb_i)$ be the bi-degrees of $f_i$. 
Suppose $g
\in (R_D)_d$ is a monomial.
 Then 
\[
	g = \prod_{i = 1}^n {f_i}^{c_i}
\]

\noindent
for some $c_i \ge - \alpha_i d$ such that $\sum_{i=1}^n c_i \bida_i = 0$
and $\sum_{i=1}^n c_i \bidb_i = 0$. We can view $g$ as an element of the
lattice 
\begin{align*}
	\Sigma = \{ (d, c_1, \ldots, c_n) \in \bz_{\geq 0}^{n+1} : c_i \ge - d
\alpha_i \text{ for all }i \text{ and }\sum_{i=1}^n c_i \bida_i = \sum_{i =
1}^n c_i \bidb_i = 0 \}.	
\end{align*}

@@

Let $\Sigma_1 \otimes_\bz \bq$ be the $\bq_{\ge 0}$ span of $e_1, \ldots e_n$.
We can intersect $\Sigma_1 \otimes_\bz \bq$ with the hyperplane $H :=
V(\sum_{i=1}^n \bida_i X_i)$  to get the subspace $\Sigma \otimes_\bz \bq = H \cap \Sigma_1
\otimes_\bz \bq$.  Then, the extremal rays of $\Sigma$ are precisely the extremal 
rays of $\Sigma \otimes_\bz \bq$.

The extremal rays of $\Sigma \otimes_\bz \bq$ can be represented by points
lying only on the edges 
$\overline{e_i e_j}$.
The extremal rays are spanned by those $e_i$'s
which are contained in $H$ together with intersection points $e_{i, j}$ which
can be expressed as $H \cap \overline{e_i e_j},$ where $i \neq j$ and $e_i,
e_j \notin H$. Note that $e_{i,j}$ is only defined in the case
$\# \{H \cap \overline{e_i, e_j}\} = 1$.

The following are elements of $R_D$ correspond to the extremal rays of
$\Sigma$.
\todo{Aaron: can we rephrase the above to say the images of the $\epsilon_i, \epsilon_{ij}$ lie on the extremal rays?}
For $i \in \Te(D)$, define $\epsilon_i \in R_D$ in degree $d_i = \ell_i \gcd(a_i, b_i)$ by 
\todo{below section needs to be simplified, maybe
by introducing new variables; 
Peter: I replaced some terms with degrees $d_i$ and $d_{i,j}$}
\begin{equation}\label{eqn:epsilon_i}
	\epsilon_i := \left(\prod_{k \ne i} f_k^{-\alpha_k d_i}\right)
	f_i^{ \frac{d_i}{\bida_i + \bidb_i}\sum_{k \ne i}
	\alpha_k (\bida_k + \bidb_k)}.
\end{equation}

@@

\noindent
and
\[
	s_2 = -\bidb_i \sum_{k \ne i, j} \alpha_k \bida_k + \bida_i \sum_{k \ne i, j}
	\alpha_k \bidb_k.
\]
\todo{Aaron: Why are $\epsilon_{i,j}$ defined here instead of $e_{i,j}$. We
don't use the $\epsilon$ here, and we do use the $e$ here. \\
Peter: I have a slight leaning towards this notation, because the algebraic
object is use more later (I think, although I'll check to make sure); also, I
think writing out the ordered tuple would actually be harder, because there is
no product notation, and it can be hard to denote the cordinates.

Aaron: I did
a search for $\epsilon$ and the last time it is mentioned is inside this proof.
In fact, while the $\epsilon$ are not referred to later, the $e_{ij}$ are referred to!
The ordered tuple will look slightly longer, but it is less confusing then
switching from tuple notation to
product notation for no reason, especially when it isn't used again.
It's confusing for the reader who is thinking in terms of lattice points
to suddently switch to thinking in a ring. At the very least, you should say 
it is the preimage of the map from the lattice to the canonical ring, since the
current use of corresponding isn't ever made clear, but
this would probably then take a bunch more space too, and I would also find this significantly more confusing than just saying the lattice points.}
A quick calculation deduces that these $\epsilon_i$'s and $\epsilon_{i,j}$'s lie on the extremal rays represented by the $e_i$'s and $e_{i,j}$'s, respectively.
\begin{rem}
For the remainder of this section replace the $e_i$'s and $e_{i,j}$'s with a (possibly) different representative of it's extremal ray say that their lattice points correspond to the $\epsilon_i$'s and $\epsilon_{i,j}$'s, respectively.
\end{rem}

\todo{Peter: I don't like having a remark inside of a proof.  Also, it feels
like this could reasonably be replaced with a better constructed/organized
proof, so that $e_i$'s are just defined this way initially; let me know if you
see a good way to do so...  As for Aaron's comment above, we could do something
like that, except we already defined $e_i$'s to be something different, which
might be equally/more confusing.
Aaron: I also don't like this here, just because it isn't really a remark,
its more of a notational convention, and it doesn't really belong in the proof.
I also think it's confusing to switch notation in the middle of a proof.
Earlier in the proof, we should just call the $e_{ij}$ something like $\hat{e_{ij}}$. Then, we should never define $\epsilon_{ij}$, since it's only used in this
proof, and should really be $e_{ij}$, as I was explaining in my previous todo.
Instead of defining $\epsilon_{ij}$, just define $e_{ij}$ which are the
appropriately scaled $\hat{e_{ij}}$ and everything will be much clearer.}
Thus,
Proposition ~\ref{prop:cone-generation} implies that 
$R_D$ is generated in 
degrees less than the sum of the degrees of the $\epsilon_i$'s and $\epsilon_{i,j}$'s, which is
\[
	\rho = \sum_{i\in \Te(D)} \gcd(\bida_i, \bidb_i)\ell_i + \sum_{\substack{
	i \in \Tp(D) \\	j \in \Tm(D)}} (\bida_i \bidb_j- \bida_j \bidb_i)\ell_{i,j}.
\]
\end{proof}

@@

\begin{lem}
\label{lem:hirz-bound-ker-psi}
$\ker \psi$ is generated in degrees less than
\[
	\tau : = \rho
	+ \max \left(\max_{i\in \Te}(\ell_i \gcd(a_i, b_i)), \max_{\substack{
	i \in	\Tp \\ j\in \Tm}} (\ell_{i,j} (\bida_i \bidb_j - \bida_j \bidb_i))
	\right).
\]
\end{lem}

\begin{proof}
We first claim that there exist $\beta_1, \ldots, \beta_n\in \bk[z_1, z_2, z_3, z_4]$
 such that $\ker \psi$ has relations of the form
\begin{equation}
\label{eqn:hirz-relations-psi}
	(z_i - \beta_i)\prod_{j=1}^n {z_j}^{c_{j}}
\end{equation}

@@

Let $\omega\in \Omega'$ correspond to the monomial $y_i \prod_{j=1}^n
{y_j}^{c_j}$.  Then we can write $\omega$ as a sum of $r_j e_j$'s for $j\in
\Te$ and $r_{j,k}e_{j,k}$ for $j\in \Tp, k\in \Tm$.  
For convenience, let $e_j$ (when it exists) be in degree $d_j$ with $j^{th}$
component $-\alpha_j d_j + \kappa_j$ for some $\kappa_j \in \bq$ and let $e_{j,k}$ (when it exists) lie in
degrees $d_{j,k}$ with $j^{th}$ component $-\alpha_j d_{j,k} + \kappa_{j,k}'$ and $k^{th}$
component $-\alpha_j d_{j,k} + \kappa_{j,k}''$ for $\kappa_{j,k}', \kappa_{j,k}'' \in \bq$
Then a careful computation deduces that

\todo{Aaron: What is $\sigma$ here? Is it $\rho$?}
%if $i\in \Te$, $i\in \Tp$, and $i\in \Tm$, then
\begin{equation}\label{eqn:sigma-in-terms-of-lambda/E}
	\sigma = \lambda + E
\end{equation}
where $\lambda \in \Sigma$ and
\begin{equation}\label{eqn:hirz-E-translation}
	E = \begin{cases}
	\frac{1}{\kappa_i} e_i  	&  \mbox{ if }i\in \Te\\
	\sum_{j\in \Tm} r_{i,j} e_{i,j}: r_{i,j}\in \bq, \sum_{j\in \Tm} r_{i,j}\kappa_{i,j}' = 1 & \mbox{ if } i\in \Tp\\
	\sum_{j\in \Tp} r_{j,i} e_{j,i}: r_{j,i} \in \bq, \sum_{j\in \Tm} r_{j,i} \kappa_{i,j}'' = 1 	& \mbox{ if } i\in \Tm\\
	\end{cases}.
\end{equation}
